Stochastic blockmodeling of relational event dynamics

### 2012-06-01 docs/experiments.tex
Space-saving stuff.

\subsection{Data}

A variety of real world data sets are used to explore the efficacy of the model.  The following data sets are sequences of dyadic events, where each event has a sender, recipient, and timestamp.  For each data set, we hold out the final $M_{test}$ events for evaluation of the model.

\begin{itemize}
%\item Kiel email: \cite{Ebel2002}
\item Classroom \cite{McFarland2001}: 445 directed communication among 27 people a high
  school classroom collected via participant observation ($M_{test}= 145$).
\item University email \cite{Eckmann2004}: 3300 dyadic emails among 88 people with at least 30
  emails each ($M_{test} = 1300$)
\item Enron email \cite{Klimt2004}: 4000 dyadic emails among 141
  individuals between July 2001 and August 2001 ($M_{test} = 1000$)
%\item Irvine: dyadic interactions among 401 actors each having more than 30 events,  \cite{Opsahl}
\item Twitter direct messages: Tweets from Twitter.com occurring between from May 11, 2009 to January 26, 2012 that contained the hashtag \texttt{\#rstats}.  This hashtag is used to denote messages pertaining to the R statistical computing environment and sometimes statistical discussion more generally.  We collect dyadic events by selecting tweets beginning with the \texttt{@} symbol (called a \emph{mention}), and mark the first mentioned user as the recipient.
Of 28337 total tweets in this time period, 3926 were directed events among a total of 1079 users.
We use a subset 4330 events among 487 users who participated in more than one event ($M_{test} = 1330$).\footnote{This dataset will be made publicly available.}
\item MIT Reality Mining \cite{Eagle2009}: 2000 phone calls among the
  89 recipients between October 2001 and February 2002 ($M_{test}= 1000$).
\end{itemize}

@@

In order to investigate the role of the number of clusters, we introduce an upper bound $K^*$ on the number of clusters during the fitting process.  In Table ~\ref{tab:results} we compare the loglikelihood on held-out test data while varying the maximum allowed $K^*$.  A relational event model (i.e. our model with $K=1$) outperforms the baselines on these data sets.  For the synthetic data example introduced in Section \ref{sec:simulation} the fitted models with $K>1$ have predictive performance comparable to the true model; a standard relational event model does not perform as well because it does not have the flexibility to model the dynamics of each block separately.  For the classroom data set we see only a slight improvement with $K>1$, while with the mobile phone calls the $K=3$ model performs best.
In Table~\ref{tab:recall5} and Table~\ref{tab:recall20} we include the corresponding results for the recall experiment at a cutoff of 5 and 20, respectively.  The results there are similar.

>>>>

%\subsection{Data}

A variety of real world data sets are used to explore the efficacy of the model.  The following data sets are sequences of dyadic events, where each event has a sender, recipient, and timestamp.  For each data set, we hold out the final $M_{test}$ events for evaluation of the model.
\begin{list}{\labelitemi}{\leftmargin=1em}
%\begin{itemize}
%\item Kiel email: \cite{Ebel2002}
\item Classroom \cite{McFarland2001}: 445 directed communication among 27 people a high
  school classroom collected via participant observation ($M_{test}= 145$).
\item University email \cite{Eckmann2004}: 3300 dyadic emails among 88
  users with at least 30 emails ($M_{test} = 1300$).
\item Enron email \cite{Klimt2004}: 4000 dyadic emails among 141
  individuals between July 2001 and August 2001 ($M_{test} = 1000$).
%\item Irvine: dyadic interactions among 401 actors each having more than 30 events,  \cite{Opsahl}
\item Twitter direct messages: Tweets from Twitter.com occurring between from May 11, 2009 to January 26, 2012 that contained the hashtag \texttt{\#rstats}.  This hashtag is used to denote messages pertaining to the R statistical computing environment and sometimes statistical discussion more generally.  We collect dyadic events by selecting tweets beginning with the \texttt{@} symbol (called a \emph{mention}), and mark the first mentioned user as the recipient.
Of 28337 total tweets in this time period, 3926 were directed events among a total of 1079 users.
We use a subset 4330 events among 487 users who participated in more than one event ($M_{test} = 1330$).\footnote{This dataset will be made publicly available.}
\item MIT Reality Mining \cite{Eagle2009}: 2000 phone calls among the
  89 recipients between October 2001 and February 2002 ($M_{test}= 1000$).
%\end{itemize}
\end{list}

@@

In order to investigate the role of the number of clusters, we introduce an upper bound $K^*$ on the number of clusters during the fitting process.  In Table ~\ref{tab:results} we compare the loglikelihood on held-out test data while varying the maximum allowed $K^*$.  A relational event model (i.e. our model with $K=1$) outperforms the baselines on these data sets.  For the synthetic data example introduced in Section \ref{sec:simulation} the fitted models with $K>1$ have predictive performance comparable to the true model; a standard relational event model does not perform as well because it does not have the flexibility to model the dynamics of each block separately.  For the classroom data set we see only a slight improvement with $K>1$, while with the mobile phone calls the $K=3$ model performs best.
In Table~\ref{tab:recall20} we include the corresponding results for the recall experiment at a cutoff of 5 and 20, respectively.  The results there are similar.
